---
title: "Assignment 1 - weather"
output: pdf_document
highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

**a)**
`r weather = read.table(file="ass1/weather.txt", header=TRUE)`


```{r, fig.height=3, fig.width=6}
hum=weather$humidity
tem=weather$temperature

par(mfrow=c(1,2))
hist(hum, main="Histogram of humidity")
hist(tem, main="Histogram of temperature")
qqnorm(hum, main="Normal Q-Q of humidity")
qqnorm(tem, main="Normal Q-Q of temperature") 
```

`r summary(hum)`
`r summary(tem)`

**b)**

```{r}
qqnorm(hum, main="Normal Q-Q of humidity")
qqnorm(tem, main="Normal Q-Q of temperature") 
```

With this plot we can see that the temperature data does not follow a pattern seen with normal distributed samples. It is not a straight line in the Q-Q plot.

**c)**

With the following R code we can calculate the 90% confidence interval.

```{r}
n = length(weather$temperature)
a = mean(weather$temperature)
s = sd(weather$temperature)
error <- qnorm(0.950) * s / sqrt(n)
low = a - margin
high = a + margin
```

This means that the true mean is (with a confidence interval of 90%) between 47.49 and 57.96.

TODO: interpretation

**d)**






